BAM's Environmental Data
=============

## Datasets

Three options to view lists of environmental / biophysical datasets used in BAM analyses. 

1. [BAM website](http://www.borealbirds.ca/index.php/biophysical_data) - Highlights the major products used up to 2012.
2. [Test Wiki](https://sites.google.com/site/bambiophysical/) describing biophysical data we have used or might use - last updated 2017, but incomplete
3. BAM's GeoNetwork data catalog -- currently **under construction**. Expected 2018. 

## Variables

### OFFSET VARIABLES

---

#### DSLS

**Explanation**: Days since local spring

**Derived from**: Spring?????

**EXAMPLE CODE**

``` {r, echo=T, eval=F}
SS_sprng <- read.csv("e:/peter/bam/May2015/NRCAN_SG_001_BAMBBS2015_71_13.csv") #load intersected file acquired from Trish
SS_sprng <- SS_sprng[,c("SS","RASTERVALU")] #subset for necessary 2 columns
SS_sprng$SPRNG <- SS_sprng$RASTERVALU 
levels(SS_sprng$SPRNG) <- gsub(",", "", levels(SS_sprng$SPRNG))
SS_sprng$SPRNG <- as.numeric(as.character(SS_sprng$SPRNG))
SS_sprng$SPRNG[SS_sprng$SPRNG < 0] <- NA
rownames(SS_sprng) <- SS_sprng$SS
SS_sprng <- SS_sprng[rownames(SS01),]

# some intermediate code not included here

DSLS <- (offdat$JULIAN - offdat$ss_sprng) / 365
```

---

#### JDAY

**Explanation**: Julian day / 365

**Derived from**: MM (month), DAY (day) -- both come from original data. If a survey doesn't have month and day, we can't calculate offsets. 

**EXAMPLE CODE**

``` {r eval=F, echo=T}
DD <- with(PKEY, paste0(YEAR, "-", MM, "-", DAY, " ", HH, ":", mm, ":00")) # combine date and time pieces back together
DD <- strptime(DD, "%Y-%m-%e %H:%M:%S") # Convert date to POSIXct.
PKEY$DATE <- DD
## Julian day
PKEY$JULIAN <- DD$yday # extract Julian (ordinal) day

PKEY$JDAY <- DD$yday / 365 # JDAY is proportion of year
```

---

#### LCC4Conif

**Explanation**: Indicator variable – if habitat is coniferous 

**Derived from**: LCC4. LCC4 is derived from [NALCMS landcover](https://sites.google.com/site/bambiophysical/home/landcover/modis-land-cover-map-of-canada-2005) 

**EXAMPLE CODE**

``` {r}
# to be filled in later
```

---

#### LCC4Open

**Explanation**: Indicator variable – if habitat is open

**Derived from**: LCC4. LCC4 is derived from [NALCMS landcover](https://sites.google.com/site/bambiophysical/home/landcover/modis-land-cover-map-of-canada-2005) 

**EXAMPLE CODE**

``` {r}
# to be filled in later
```

---

#### LCC2OpenWet

**Explanation**: Indicator variable - if habitat is open or wet

**Derived from**: LCC2. LCC2 is derived from [NALCMS landcover](https://sites.google.com/site/bambiophysical/home/landcover/modis-land-cover-map-of-canada-2005) 

**EXAMPLE CODE**

``` {r}
# to be filled in later
```

---

#### LCC4Wet
**Explanation**: Indicator variable – if habitat is wet (from LCC4)

**Derived from**: LCC2. LCC2 is derived from [NALCMS landcover](https://sites.google.com/site/bambiophysical/home/landcover/modis-land-cover-map-of-canada-2005) 

**EXAMPLE CODE**

``` {r}
# to be filled in later
```

---

#### MAXDIS

**Explanation**: Maximum distance band for point count in m, divided by 100

**Derived from**: MASDIST, joined to a PKEY via METHOD. METHOD supplied in raw survey methodology

**EXAMPLE CODE**

``` {r}
# to be filled in later
```

---

#### TREE

**Explanation**: Percent tree cover

**Derived from**: Intersection with a tree cover layer -- which one??

**EXAMPLE CODE**

``` {r}
# to be filled in later
```

#### TSSR

**Explanation**: Time since sunrise (requires time zone)

**Derived from**: HH (hour), MIN (minute), time zone

**EXAMPLE CODE**

``` {r}
# to be filled in later
```


### LANDCOVER VARIABLES

#### LCC05

**Explanation**: 

**Derived from**: 

**EXAMPLE CODE**

``` {r}
# to be filled in later
```

---

#### LCC2

**Explanation**: 

**Derived from**: 

**EXAMPLE CODE**

``` {r}
# to be filled in later
```

---

#### LCC4

**Explanation**: 

**Derived from**: 

**EXAMPLE CODE**

``` {r}
# to be filled in later
```


### METHODOLOGICAL VARIABLES

#### ROAD

**Explanation**: 

**Derived from**: 

**EXAMPLE CODE**

``` {r}
# to be filled in later
```

---



